---
title: "1. Evaluation of a Pascal Compiler"
author: "Robert A. Stevens"
date: "8/23/2017"
output: html_document
---

```{r, comment=NA}

```

*Pascal Programs for Scientists and Engineers*

by Alan R. Miller

## 1.1 Introduction

## 1.2 Precision and Range of Floating-Point Operations

## 1.3 Pascal Program: A Test of the Floating-Point Operations

http://infohost.nmt.edu/~es421/pascal/list11.pas

```
PROGRAM Test;
 { test range of floating-point numbers }
 { From Borland Pascal Programs for Scientists and Engineers }
 { by Alan R. Miller, Copyright C 1993, SYBEX Inc }

USES Crt; { Crt for non-windows version}

VAR
  I: Integer;
  X: Real;

BEGIN
  WriteLn;
  X := 1.0E-4 / 3.0;
  FOR I := 1 TO 18 DO
    BEGIN
      Write( I:5, X);
      X := 0.1 * X;
      WriteLn('    ', X);
      X := 0.1 * X
    END;
  Writeln;
  Writeln('  Press Enter to end');
  REPEAT UNTIL KeyPressed;
END.
```

```{r, comment=NA}
X <- 1.0E-4 / 3.0
for(i in 1:18) {
  cat(i, X);
  X <- 0.1 * X
  cat('    ', X, '\n')
  X <- 0.1 * X
}

# without "for" loop
X <- (1/3)/10^(4:39)
matrix(X, ncol = 2, byrow = TRUE)
```

**Figure 1.1:** A Test of the Floating-Point Operations

**Figure 1.2:** Precision Test: First Complier

**Figure 1.3:** Precision Test: Second Complier

**Figure 1.4:** Precision Test: Third Complier

## 1.4 Pascal SIN and COS Functions

## 1.5 Pascal Program: Testing the SIN Function

```
PROGRAM TSIN;
 { test range of sin }

USES WinCrt; { Crt for non-windows version}

VAR
  I: Integer;
  X, Y: Real;

BEGIN
  X := 1.0E-4 / 0.3;
  FOR I := 1 TO 40 DO
    BEGIN
      Y := SIN(X);
      WriteLn(' X =', X,', sin =', Y);
      X := 0.1 * X
    END;
  Writeln; Writeln('  Press Enter to end');
  REPEAT UNTIL KeyPressed; 
  DoneWinCrt  { for Windows version only }
END.
```

```{r, comment=NA}
#X <- 1.0E-4 / 3.0
X <- 1.0E-3 / 3.0
for(i in 1:15) {
  cat(i, X, '    ', sin(X), '\n')
  X <- 0.1 * X
}

# without "for" loop
X <- (1/3)/10^(3:17)
Y <- sin(X)
cbind(X, Y)
```

**Figure 1.5:** Test for the SIN Function

**Figure 1.6:** SIN Test: First Complier

**Figure 1.7:** SIN Test: Second Complier

## 1.6 Other Pascal Functions

http://infohost.nmt.edu/~es421/pascal/list12.pas

```
PROGRAM Tlog;
 { test ln and exp }
 { From Borland Pascal Programs for Scientists and Engineers }
 { by Alan R. Miller, Copyright C 1993, SYBEX Inc }
USES WinCrt; { Crt for non-windows version}

VAR
  I: Integer;
  X, Y: Real;

BEGIN
  X := 1.0E-4 / 0.3;
  FOR I := 1 TO 10 DO
    BEGIN
      Y := Ln(X);
      WriteLn(' X =', X,', Exp(Ln) =', Exp(Y));
      X := 0.5 * X
    END;
  Writeln; Writeln('  Press Enter to end');
  REPEAT UNTIL KeyPressed; 
  DoneWinCrt  { for Windows version only }
END.
```

```{r, comment=NA}
X <- 1.0E-4 / 0.3
for (i in 1:10) {
  Y <- log(X)
  cat(' X =', X, ', Exp(Ln) =', exp(Y), "\n")
  X <- 0.5 * X
}

# without "for" loop
X <- (0.5^c(0:9))/(3*10^3)
Y <- log(X)
LabelX <- rep(" X =", 10)
LabelY <- rep(", Exp(Ln) =", 10)
data.frame(LabelX, X, LabelY, expY = exp(Y))
```

**Figure 1.8:** A Test of the LN and EXP Functions

```{r, comment=NA}

```

**Figure 1.9:** An Arc Tangent Function with Two Arguments

```{r, comment=NA}

```

**Figure 1.10:** The Arc Sine Function

```{r, comment=NA}

```

**Figure 1.11:** The Arc Cosine Function

```{r, comment=NA}

```

**Figure 1.12:** Program to Test the Arc Sine and Arc Tangent Functions

## 1.7 External Files

## 1.8 A Power-of-10 Function

## 1.9 Pascal Program: Calculating Powers of 10

```{r, comment=NA}

```

**Figure 1.13:** A Function to Calculate Powers of 10

```{r, comment=NA}

```

**Figure 1.14:** Output from the Power-of-10 Program

## 1.10 Summary
